Prototype city generation for mobility simulation in urban typologies: a generalized framework

### 2017-12-15 manuscript/manuscript.tex
revision to framework document

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{listings}
\usepackage[english]{babel}
\usepackage{enumerate}
\lstset{frame=tlrb,xleftmargin=\fboxsep,xrightmargin=-\fboxsep}

\usepackage[style=numeric, backend=biber, maxcitenames=2]{biblatex}
\bibliography{bib/references}

\begin{document}

\title{Population Synthesis Literature Review}
\author{Jin Zhou}
\maketitle

\section{Goal}
Since the step of urban typology has been done, we need to fill in the blank between city clustering results and a complete SimMobility\cite{adnan2016simmobility} input of virtual city.
The city model that SimMobility takes as input has two components.
One is the demand and the other is the supply, which includes the road and transit network.
To model the demand, a methodology has to be developed to construct the population with desired demographics and allocate the population in space.
Next to it, we also need to figure out how to use the synthesized households and individuals in disaggregate level to get the travel demand in the city.
The literature review at this time is to gain a better idea of previous efforts to address the problem of population synthesis and assignment.

\subsection{Questions:}
\begin{enumerate}
	\item What should the input of travel demand generation look like? This guides the adaptation of population synthesis method from literature.
	\item What attributes should be included to characterize the population? Age, sex, educational level, and there must be some attributes reflect the traveling behavior of people. 
	\item Do we synthesize households or personnel? Or how to match individuals to households?
	\item We need to come up with a solution to cities don't have PUMS data and virtual cities that don't even have census data for demographics.	
\end{enumerate}

\section{Prior Work}

Household assignment and spatial allocation are two big issues we need to solve when facing the problem of population synthesis.
Specifically, the first one addresses the problem of how to match individuals to the household where they belong, while the spatial allocation distributes households in geographical space.
It is necessary because the ultimate goal of our work is to do transportation simulation, which can only be achieved by knowing the location of each agent. \\

There has been lots of literature on population synthesis.
The methods used varies from Iterative Proportional Fitting (IPF), MCMC-based sampling \cite{farooq2013simulation}, to Bayesian network approach \cite{sun2015bayesian}. The original work in this field is presented in \cite{beckman1996creating}. It used a method called Iterative Proportional Fitting (IPF) to synthesize population based on data from two sources. What they do is iteratively fitting the cells of a multi-way contingency table expanded by attributes that characterize a household/individual so that it satisfies the constraints of marginal distribution of demographics and also preserves the correlation structure carried by sample data of the whole population. Here the marginals of demographics comes from census data in the level of census tract or block group while the correlation structure need to be derived from survey data of target population. In case of the absence of disaggregated sample data, IPF can still give a solution by iteratively adjust the cells in accordance only with the marginal constraints. However, lots of different joint distributions can share the same marginals and IPF may converge to anyone of them. This is certainly a downside of traditional IPF, and is also the reason why we need the correlation structure of sample data to make the solution captures the true population distribution. After the fitting of the table, a desired number of synthetic population is sampled according to individual or household weights in the table. And this is where the sensitivity of IPF to the quality of data and the sample size stems from. In this method, spatial allocation is solved concurrently with the population synthesis, since it generates the population for each zone and then adds up to the whole population. On contrary to this, one can also synthesize the total population first and then allocate them to different zones. However, this includes implementing a detailed matching method between households and zones, like what's developed in \cite{ge2014virtual}, which is also very time-consuming.

@@

The population synthesis method that \textcite{petrik2016measuring} used adapted ideas from probabilistic graphical models (PGM).
In their graphical model, there are 7 socio-demographic attributes, which are Age, Gender, Employment, Income, Car ownership, Purpose of trip, and Who pay for the trip, to represent a person.
Once the dependencies between variables are specified, we need to calculate or estimate conditional probability of one node given its parents.
The flexibility of this method is that those conditional probabilities can be derived from different sources like census data or surveys on individuals.
Finally, the synthetic population is just random draws from the probabilistic graphical model.
Though we can take marginal distributions of all variables as supplement for the graphical model, the population generated by Bayesian network model can only precisely match the marginals of mother nodes.
This is because the distribution of a node is uniquely defined by its parent node and the conditional distribution based on PGM.
And this point should be taken into consideration when choosing the population synthesis method.

\section{Data}
Most paper in population synthesis used the Public Use Microdata Sample (PUMS), which is available both in individual level and household level.
It includes multiple socio-economic parameters of a household or an individual in the region.
The sample rate of this dataset is about 1\% to 10\% of the total population in the region.
We found PUMS data is complete in US cities but haven't found microdata in other cities in the world.
Currently, we have found aggregated information of population, age, sex, household, education in districts in Beijing.

The choice of algorithm should be based both on the data we have and the appearance of the final result we want.
What is also important for us is the simplicity and novelty of methodology.

\section{Iterative Proportional Fitting}

Using IPF to synthesize population includes two major steps.
In the first step, the disaggregate samples of population are used to initialize the contingency table expanded by attributes, and the fitting is carried out to meet marginal constraints from aggregated data.
In the next step, the fitted contingency table is used to sample the population.
To describe IPF, we need to define the contingency table it fits.
Suppose a person or household is described by \textit{m} attributes (or demographics), then we need to develop a m-way table.
For each demographic $i$, there are $n_i$ categories.
Thus, a cell $(i_1, i_2,..., i_m)$ represents a kind of person/household, where $i_j = 1,2,...n_j$ is the observed value of the $j$-th demographic with $n_j$ categories it can choose from.
From sample data like PUMS, $n$ is the total number of observations while $n_{i_1, i_2,...,i_m}$ is the counts for people/household of this type.
Hence, the proportion of this type in the sample is denoted by
\begin{equation}
  \label{eqn:prop}
   p_{i_1,i_2,...,i_m} = \frac{n_{i_1, i_2,...,i_m}}{n}
\end{equation}
The constraints is defined as $T_k^j$, which is the marginal totals for the $k$-th category of attribute $i$ from census data.
We can easily get the equation
\begin{equation}
  \label{eq:2}
\sum_i\sum_{k=1}^{n_i}T_k^j=n.  
\end{equation}
Let $p_{i_1,i_2,...,i_m}^{(t)}$ denote estimated proportion of cell $(i_1, i_2,..., i_m)$ in iteration $t$ and
\begin{equation}
  \label{eq:3}
  p_{...,i_j=k,...} = \sum_{i_1=1}^{n_1}...\sum_{i_m=1}^{n_m}p_{i_1,i_2,...,i_j=k,...,i_m}^{(t)}  
\end{equation}

In the next step, the synthetic population shall be constructed by selecting households from PUMS in proportion to the estimated weights in the multiway table.
Specifically, the number of households of each demographic type in a census tract can be obtained by multiplying the total number of households by the probability in the cell representing this household type, or by drawing the numbers at random according to these probabilities.
However, the household type represent by a cell in multiway table is not exactly that of PUMS data.
The fitted multiway table can only be span by control variables, which is defined as variables which have marginal distribution in census data.
However, households and individuals in disaggregate data are usually described by additional desired variables.
So, the probability of a household type in PUMS data being chosen shall be assigned according to the distance between such a household $p$ and a household type $c$ in multi-way table.
In \cite{beckman1996creating}, this probability is defined as 
\begin{equation}
	D(p,c) = w_p \prod_{i \in J}\Big(1-|(d_i^p-d_i^c)/r_i|^k\Big)*\prod_{i \notin J}\Big(1-\Delta(d_i^p, d_i^c)\Big)
\end{equation}

\section{Proposed method}

\begin{enumerate}[Step 1:]
\item Network: modify OSM network file, make it legal input to SimMobility
  \begin{enumerate}[1.]
  \item links has to be unidirectional
  \item shorten the links and add turning points
  \item maintain the attributes of each object in the network
  \end{enumerate}
\item Population Synthesis
  \begin{enumerate}[1. ]
  \item Data collection: survey data, aggregate census data for zones in each city
  \item Learn the parameter and the structure of Bayesian Network which captures the structure of population
  \item sampling from Bayesian Network to get whole population in the city
  \item Clustering the synthetic population in a city by K-means
  \item Divide the city into zones according to clustering result, and form the aggregate census data for zones
  \end{enumerate} 
\end{enumerate}

@@

\subsection{Two approaches to spatial assignment of households}
\subsubsection{K-means clustering}
With regard to the assignment of populations to zones, we will cluster the population at city level with K-means first.
We reason that similar and highly correlated people will most likely live in the same area within a city.
Then we obtain grouped people with similar attributes.
We choose K-means because it is more intuitive if we want a spatial allocation and the feature of hidden centroid for each cluster will hopefully capture the characteristic of the zone.

@@

\newpage
\section{References}
\printbibliography[heading=none]
\end{document}

>>>>

\documentclass[11pt,twoside]{article}
\usepackage{etex}
\newcommand{\num}{6{} }

\raggedbottom

%geometry (sets margin) and other useful packages
\usepackage{geometry}
\geometry{top=1in, left=1in,right=1in,bottom=1in}
 \usepackage{graphicx,booktabs,calc}

%=== GRAPHICS PATH ===========
\graphicspath{{./images/}}
% Marginpar width
%Marginpar width
\newcommand{\pts}[1]{\marginpar{ \small\hspace{0pt} \textit{[#1]} } } 
\setlength{\marginparwidth}{.5in}
%\reversemarginpar
%\setlength{\marginparsep}{.02in}

%% Fonts
% \usepackage{fourier}
% \usepackage[T1]{pbsi}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
%\usepackage{minted}

\usepackage{rotating}
%% Cite Title
\usepackage[style=authoryear,maxcitenames=2,natbib]{biblatex}
\addbibresource{../bib/references.bib}

%%% Counters
\usepackage{chngcntr,mathtools}
\counterwithout{figure}{section}
\counterwithout{table}{section}

\numberwithin{equation}{section}

%% Captions
\usepackage{caption}
\captionsetup{
  labelsep=quad,
  justification=raggedright,
  labelfont=sc
}

%AMS-TeX packages
\usepackage{amssymb,amsmath,amsthm} 
\usepackage{bm}
\usepackage[mathscr]{eucal}
\usepackage{colortbl}
\usepackage{color}

\usepackage{epstopdf,subfigure,hyperref,enumerate,polynom,polynomial}
\usepackage{multirow,minitoc,fancybox,array,multicol}

\definecolor{slblue}{rgb}{0,.3,.62}%
%\renewcommand{\chaptermark}[1]{ \markboth{#1}{} }
\renewcommand{\sectionmark}[1]{ \markright{#1}{} }

\usepackage{fancyhdr}
\pagestyle{fancy}
%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
%\addtolength{\headwidth}{\marginparwidth}
%\fancyheadoffset{30pt}
%\fancyfootoffset{30pt}
\fancyhead[LO,RE]{\small  \it \nouppercase{\leftmark}}
\fancyhead[RO,LE]{\small Page \thepage} 
\fancyfoot[RO,LE]{\small }% PR \num S-2015} 
\fancyfoot[LO,RE]{\small }%\scshape MODL} 
\cfoot{} 
\renewcommand{\headrulewidth}{0.1pt} 
\renewcommand{\footrulewidth}{0pt}
%\setlength\voffset{-0.25in}
%\setlength\textheight{648pt}

\usepackage{paralist}

%%% FORMAT PYTHON CODE
\usepackage{listings}
% Default fixed font does not support bold face
\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{8} % for bold
\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{8}  % for normal

% Custom colors
\usepackage{color}
\definecolor{deepblue}{rgb}{0,0,0.5}
\definecolor{deepred}{rgb}{0.6,0,0}
\definecolor{deepgreen}{rgb}{0,0.5,0}

%\usepackage{listings}

% % Python style for highlighting
% \newcommand\pythonstyle{\lstset{
% language=Python,
% basicstyle=\footnotesize\ttm,
% otherkeywords={self},             % Add keywords here
% keywordstyle=\footnotesize\ttb\color{deepblue},
% emph={MyClass,__init__},          % Custom highlighting
% emphstyle=\footnotesize\ttb\color{deepred},    % Custom highlighting style
% stringstyle=\color{deepgreen},
% frame=tb,                         % Any extra options here
% showstringspaces=false            % 
% }}

% % Python environment
% \lstnewenvironment{python}[1][]
% {
% \pythonstyle
% \lstset{#1}
% }
% {}

% % Python for external files
% \newcommand\pythonexternal[2][]{{
% \pythonstyle
% \lstinputlisting[#1]{#2}}}

% % Python for inline
% \newcommand\pythoninline[1]{{\pythonstyle\lstinline!#1!}}

\newcommand{\osn}{\oldstylenums}
\newcommand{\dg}{^{\circ}}
\newcommand{\lt}{\left}
\newcommand{\rt}{\right}
\newcommand{\pt}{\phantom}
\newcommand{\tf}{\therefore}
\newcommand{\?}{\stackrel{?}{=}}
\newcommand{\fr}{\frac}
\newcommand{\dfr}{\dfrac}
\newcommand{\ul}{\underline}
\newcommand{\tn}{\tabularnewline}
\newcommand{\nl}{\newline}
\newcommand\relph[1]{\mathrel{\phantom{#1}}}
\newcommand{\cm}{\checkmark}
\newcommand{\ol}{\overline}
\newcommand{\rd}{\color{red}}
\newcommand{\bl}{\color{blue}}
\newcommand{\pl}{\color{purple}}
\newcommand{\og}{\color{orange!90!black}}
\newcommand{\gr}{\color{green!40!black}}
\newcommand{\nin}{\noindent}
\newcommand{\la}{\lambda}
\renewcommand{\th}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\G}{\Gamma}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,thick,inner sep=1pt] (char) {\small #1};}}

\newcommand{\bc}{\begin{compactenum}[\quad--]}
\newcommand{\ec}{\end{compactenum}}

\newcommand{\p}{\partial}
\newcommand{\pd}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\dpd}[2]{\dfrac{\partial{#1}}{\partial{#2}}}
\newcommand{\pdd}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Prototype city generation for mobility simulation in urban typologies: a  generalized framework}
\author{}
\maketitle

\section{Introduction}
Household assignment and spatial allocation are two big issues we need to solve when facing the problem of population synthesis.
Specifically, the first one addresses the problem of how to match individuals to the household where they belong, while the spatial allocation distributes households in geographical space.
It is necessary because the ultimate goal of our work is to do transportation simulation, which can only be achieved by knowing the location of each agent. \\

\section{Background}
There has been lots of literature on population synthesis.
The methods used varies from Iterative Proportional Fitting (IPF), MCMC-based sampling \cite{farooq2013simulation}, to Bayesian network approach \cite{sun2015bayesian}. The original work in this field is presented in \cite{beckman1996creating}. It used a method called Iterative Proportional Fitting (IPF) to synthesize population based on data from two sources. What they do is iteratively fitting the cells of a multi-way contingency table expanded by attributes that characterize a household/individual so that it satisfies the constraints of marginal distribution of demographics and also preserves the correlation structure carried by sample data of the whole population. Here the marginals of demographics comes from census data in the level of census tract or block group while the correlation structure need to be derived from survey data of target population. In case of the absence of disaggregated sample data, IPF can still give a solution by iteratively adjust the cells in accordance only with the marginal constraints. However, lots of different joint distributions can share the same marginals and IPF may converge to anyone of them. This is certainly a downside of traditional IPF, and is also the reason why we need the correlation structure of sample data to make the solution captures the true population distribution. After the fitting of the table, a desired number of synthetic population is sampled according to individual or household weights in the table. And this is where the sensitivity of IPF to the quality of data and the sample size stems from. In this method, spatial allocation is solved concurrently with the population synthesis, since it generates the population for each zone and then adds up to the whole population. On contrary to this, one can also synthesize the total population first and then allocate them to different zones. However, this includes implementing a detailed matching method between households and zones, like what's developed in \cite{ge2014virtual}, which is also very time-consuming.

@@

The population synthesis method that %\cite{petrik2016measuring}
used adapted ideas from probabilistic graphical models (PGM).
In their graphical model, there are 7 socio-demographic attributes, which are Age, Gender, Employment, Income, Car ownership, Purpose of trip, and Who pay for the trip, to represent a person.
Once the dependencies between variables are specified, we need to calculate or estimate conditional probability of one node given its parents.
The flexibility of this method is that those conditional probabilities can be derived from different sources like census data or surveys on individuals. Nevertheless, compared to the high dependency of IPF on the quality of sample data, Bayesian method gives synthetic population more heterogeneity.
Finally, the synthetic population is just random draws from the probabilistic graphical model.
Though we can take marginal distributions of all variables as supplement for the graphical model, the population generated by Bayesian network model can only precisely match the marginals of mother nodes.
This is because the distribution of a node is uniquely defined by its parent node and the conditional distribution based on PGM.
And this point should be taken into consideration when choosing the population synthesis method.

\section{Data}
Most paper in population synthesis used Public Use Microdata Sample (PUMS) data, which is available both in individual level and household level.
It includes multiple socio-economic parameters of a household or an individual in the region.
The sample rate of this dataset is about 1\% to 10\% of the total population in the region. 
PUMS data is very complete in US cities and can be easily downloaded from government census website \footnote{https://www.census.gov/programs-surveys/acs/data/pums.html}, but the same data for cities in developing countries can be relatively outdated and harder to find. The sample data is used as seeds to initialize the contingency table in IPF methods. Another important part of data is called aggregate data \footnote{https://factfinder.census.gov}. They give marginal distribution for variables of our interests. It's worth noticing that the marginal distributions for subsets with more than 2 variables can be rarely found. An ideal format for aggregate data is that we have the marginal distributions in the level of geographical zone cause IPF is carried out separately in each zone. So when downloading the aggregate data, one should first choose a geographic type in corresponding to the zones in the virtual city, then specify the variables or tables you want.

The choice of algorithm should be based both on the data we have and the appearance of the final result we want.
What is also important for us is the simplicity and novelty of methodology. We will start from a conventional IPF method using data from Houston city, then according to the execution time and the accuracy of the result we will make further extension to IPF method.

\subsection{Network processing}

\section{Method}

\begin{table}[h!]
  \centering
  \begin{tabular}{l l l}
    &&
  \end{tabular}
  \caption{List of symbols for variables and parameters}
  \label{tab:symbols}
\end{table}

\begin{enumerate}[Step 1:]
\item Network: modify OSM network file, make it legal input to SimMobility
  \begin{enumerate}[1.]
  \item links has to be unidirectional
  \item shorten the links and add turning points
  \item maintain the attributes of each object in the network
  \end{enumerate}
\item Population Synthesis
  \begin{enumerate}[1.]
  \item Data collection: sample data, aggregate data for zones in each city
  \item Data processing: categorize sample data and get joint distribution, formatting aggregate data
  \item Apply IPF algorithm and use the weights result to draw synthetic population
  \item Assign households in each county to locations based on land use information
  \end{enumerate} 
\end{enumerate}

\subsection{Supply}

\subsection{Demand}

@@

@@

\subsection{Summary of framework}
In our first exploration, we have used 5 attributes to describe a households and they are household type, household size, household income in the last 12 months, number of vehicles and number of workers in the household. We not only find one dimensional marginal distribution for those variables but also two dimensional marginal distributions for household type $\times$ household size, number of workers $\times$ household size as well as number of vehicles $\times$ number of workers in the household. Across the process of data retrieving and processing, we find the categories each variable can have is largely depends on the availability and format of aggregate data.

\section{Results and discussion}

\section{Conclusion}
\section{References}
\printbibliography

\appendix
\section{Further notes on hierarchical iterative proportional fitting}
Using IPF to synthesize population includes two major steps.
In the first step, the disaggregate samples of population are used to initialize the contingency table expanded by attributes, and the fitting is carried out to meet marginal constraints from aggregated data.
In the next step, the fitted contingency table is used to sample the population.
To describe IPF, we need to define the contingency table it fits.
Suppose a person or household is described by \textit{m} attributes (or demographics), then we need to develop a m-way table.
For each demographic $i$, there are $n_i$ categories.
Thus, a cell $(i_1, i_2,..., i_m)$ represents a kind of person/household, where $i_j = 1,2,...n_j$ is the observed value of the $j$-th demographic with $n_j$ categories it can choose from.
From sample data like PUMS, $n$ is the total number of observations while $n_{i_1, i_2,...,i_m}$ is the counts for people/household of this type.
Hence, the proportion of this type in the sample is denoted by
\begin{equation}
  \label{eqn:prop}
   p_{i_1,i_2,...,i_m} = \frac{n_{i_1, i_2,...,i_m}}{n}
\end{equation}
The constraints is defined as $T_k^j$, which is the marginal totals for the $k$-th category of attribute $i$ from census data.
We can easily get the equation
\begin{equation}
  \label{eq:2}
\sum_i\sum_{k=1}^{n_i}T_k^j=n.  
\end{equation}
Let $p_{i_1,i_2,...,i_m}^{(t)}$ denote estimated proportion of cell $(i_1, i_2,..., i_m)$ in iteration $t$ and
\begin{equation}
  \label{eq:3}
  p_{...,i_j=k,...} = \sum_{i_1=1}^{n_1}...\sum_{i_m=1}^{n_m}p_{i_1,i_2,...,i_j=k,...,i_m}^{(t)}  
\end{equation}

In the next step, the synthetic population shall be constructed by selecting households from PUMS in proportion to the estimated weights in the multiway table. Or we can do Monte Carlo sampling method to draw households from estimated weights from IPF.
Specifically, the number of households of each demographic type in a census tract can be obtained by multiplying the total number of households by the probability in the cell representing this household type, or by drawing the numbers at random according to these probabilities.
However, the household type represent by a cell in multiway table is not exactly that of PUMS data.
The fitted multiway table can only be span by control variables, which is defined as variables which have marginal distribution in census data.
However, households and individuals in disaggregate data are usually described by additional desired variables.
So, the probability of a household type in PUMS data being chosen shall be assigned according to the distance between such a household $p$ and a household type $c$ in multi-way table.
In \cite{beckman1996creating}, this probability is defined as 
\begin{equation}
  D(p,c) = w_p \prod_{i \in J}\Big(1-|(d_i^p-d_i^c)/r_i|^k\Big)*\prod_{i \notin J}\Big(1-\Delta(d_i^p, d_i^c)\Big)
\end{equation}

\end{document}
